Static Discharge of Dynamic, Higher-Order Contracts

### 2013-11-16 pldi2014/content/discharge.tex
Everything but discharge section.

We are after a similar result for chaperoned procedures which will allow us to identify meaningful chaperoned values---values which influence the behavior of the program via chaperoning.
While chaperoned values do so primarily by raising an error, chaperoned procedures can affect program behavior by diverging.
While we expect that programs will seldom result in divergence, and that specifically at the hand of a chaperoned value, specifications expressed in the source language are susceptible to the same programmer errors as the program proper, which includes unintended partiality.

@@

If a specific chaperoned value is not meaningful in a particular execution, it can be soundly \emph{discharged}, meaning that it is safe to replace its particular installation with identity if it is possible to do so.
(The soundness claim does not say it is necessarily possible---only safe---to do so.)
A chaperone is meaningful if
\begin{enumerate}
\item the negative guard causes an error to be raised,
\item \label{con:neg-divergence} the negative guard diverges,
\item the negative guard does not return the same number of results as arguments given (or one extra),
\item the extra result, if provided, is not a procedure,
\item the results of the negative guard are not, at most, chaperones of the arguments,
\item the positive guard does not accept the number of results produced by the chaperoned procedure,
\item the positive guard causes an error to be raised,
\item the positive guard diverges,
\item the positive guard does not return the same number of results as arguments given,
\item the results of the positive guard are not, at most, chaperones of the arguments, or
\item its application effects the installation of a meaningful chaperone.
\end{enumerate}

@@

This strategy immediately shifts our focus from dynamic points of installation to static points of installation in the following way:
First, chaperone installation may not be the only operation performed at a call site.
Thus, we must restrict ourselves to call sites dedicated to that purpose.
Second, even once we have restricted ourselves to bonafide chaperone-installation sites, there may be more than one chaperone installed at each site.
Thus, we must restrict ourselves to sites at which no meaningful chaperone is installed.
Finally, we must extend these conditions from a particular execution to all possible executions.
% multiple operations performed at a particular site
% multiple chaperones installed at a particular site (during one particular execution)
% the chaper
First, multiple chaperone installations may occur at a given application site during program execution, so we cannot, in all circumstances, selectively prevent the creation of a single chaperone.

to points of installation

This strategy immediately raises three issues.
Such a strategy affects all executions

Using this approach to discharge, it is clear that chaperone installment at a particular site can be discharged only if it meets these conditions for every possible execution.
These conditions seem onerous.
Indeed, in Racket, users typically access chaperone installation through a library wrapper so all chaperones in the entire program share the same installation site.
This, in turn, means that a given chaperone can be safely discharged only if \emph{all} chaperones in all executions of the program can be safely discharged.
%In other words, the program must be correct up to the encoded specification.
Fortunately, with access to the whole program, we may inline definitions which allows us to multiply the number of installment sites in the program.

This immediately raises two issues.

Second, chaperone installation may not be the only operation performed by an application site.
We address both by analyzing application sites and then only those in which \scheme{chap-op} is the only operator, which we refer to as \emph{chaperone-installation sites}.
The conditions to discharge a given site are stronger than that of a chaperone installed at that site: instead of demonstrating that a single chaperone is not meaningful, we must demonstrate that every chaperone created at that site is not meaningful \emph{and} that the chaperone installations themselves are not meaningful.

@@

The conditions necessary for the sound discharge of a chaperone are more easily constructed by considering the conditions sufficient for a chaperone to be meaningful.
If we can enumerate a complete set of these conditions in the sense that any meaningful chaperone must meet at least one of the conditions, we obtain the soundness conditions by complement.

For a given program execution, if neither a chaperone nor its installation raises an error or diverges, it is a candidate for \emph{discharge}, or the targeted prevention of its installation.
It is only a candidate because its application may cause a meaningful chaperone to be installed, i.e., a chaperone that is not so well-behaved.
In this case, discharging the original chaperone would prevent a meaningful chaperone from being created which would make discharge unsound.
Then, intuitively, a chaperone can be discharged if it will not raise an error or diverge  \emph{and} any chaperones it installs can be discharged.

@@

Much of the run-time overhead of chaperones is shared by impersonators.
Nevertheless, the enforcement of invariants specific to chaperones is somewhat costly --- specifically, the enforcement that chaperone guards at most install chaperones on their arguments.
The conditions to \emph{demote} the set of chaperones installed at a particular site to impersonators (by replacing the \scheme{chap-op} operator with \scheme{imp-op}) are immediately weaker than to discharge the site entirely: we need to demonstrate only that the guards of chaperones installed there do not arbitrarily modify their arguments (conditions 5 and 10 above) for it to be a candidate for demotion.
If one of a site's chaperones is installed on the argument of a chaperone guard, demoting it to install merely an impersonator will make an otherwise well-behaved chaperone not so.

For instance, the program
\begin{schemedisplay}
(let ([sabotage (chap-op (LAMBDA (f) f) (LAMBDA (f) (chap-op f raise)))])
  (let ([g (sabotage (LAMBDA (x) x))])
    (g 1)))
\end{schemedisplay}
defines a chaperoned procedure \scheme{sabotage} which wraps higher-order arguments to fail on application, defines \scheme{g} by applying \scheme{sabotage} to the identity function, and applies \scheme{g} to \scheme{1}, which raises an error.
At no point does \scheme{g} violate chaperone-specific restrictions, so it is tempting to alter its installation site (within the guard of \scheme{sabotage}) to merely impersonate the first argument.
If done, however, the program would fail at the earlier point when \scheme{sabotage} is called which certainly differs in meaning.

This observation provides us with the intuition behind sound site demotion: a chaperone may be demoted if neither of its guards do more than install chaperones on their arguments for all program executions and the chaperone which guard installed it, if any, can be and is demoted. 
We once again lift the conditions to chaperone-installment sites so that a chaperone-installment site can be demoted if all chaperones installed there in any program execution can be demoted and the installment sites of any installing chaperones can be and are demoted.

@@

We reverse the notion of dependence to say that certain control events depend on the active procedures on the stack.
With our ultimate goal being to prevent chaperone installation, we are only concerned with the chaperone guard procedures depended upon at each point of interest.
Even in the presence of tail-call optimization, as in \chapcalc, the presence of active chaperone guard procedures is always recorded on the stack, and we can apply Harrison's dependence reasoning fairly directly.

@@

Let $H_c(c_{\ell_1},c_{\ell_0})$ if $c_{\ell_1}$ depends on $c_{\ell_0}$.

@@

Let $H_s(s_{\ell_1},c_{\ell_0})$ if $s_{\ell_1}$ depends on $c_{\ell_0}$.

@@

Let $H_e(e,c_\ell)$ if $e$ depends on $c_\ell$.

@@

Let $H_e(e,s_\ell)$ if $e$ depends on $s_\ell$.

@@

Let $G(c_\ell)$ hold if $c_\ell$ is well-behaved.

@@

Transitive, reflexive closure of $H_c$ is $\bar{H_c}$,
then $c_{\ell_0}$ is dischargeable if for all $c_{\ell_1}$ such that $\bar{H_c}(c_{\ell_1},c_{\ell_0})$, $G(c_{\ell_1})$.

@@

\begin{proof}
By induction on the reduction relation.
Suppose $p\rrs v$.
Then none of the chaperone guards (or any guard, for that matter) must raise an error.
Then there is nothing to worry about.
That is, $D_\ell(p)\rrs v$, where we make no distinction between an operator and a chaperone of it.

Suppose $p\Uparrow b$. 
Then an error must be raised while a chaperone (with $b$ derived from its label) is active.
Then that chaperone depends on the error.
Then that chaperone-installment site will not be a candidate for deletion.
It is not dischargeable.
Then it's not labelled $\ell$.
Since this is true for every dependent non-dischargeable site (that it's not labelled $\ell$), then $D_\ell(p)\Uparrow b$.

Suppose $p\Uparrow$.
If the installation or application of a chaperone could cause divergence, it won't be dischargeable.
Then the chaperone site labelled $\ell$ must not cause divergence.
Then $D_\ell(p)\Uparrow$.
\end{proof}

@@

\begin{proof}
Suppose $p\rrs v$.
Then no error is thrown in regular execution.
Because we will not demote a site unless there is no possibility that its result is used by a chaperone, demoting the site will not introduce an error.
Therefore, $E_\ell(p)\rrs v$ where we make no distinction between an operator and an impersonator of it.

\end{proof}

@@

\begin{proof}
Suppose $p\rrs v$.
Then no error is thrown in regular execution.
Because we will not demote a site unless there is no possibility that its result is used by a chaperone, demoting the site will not introduce an error.
Therefore, $E_\ell(p)\rrs v$ where we make no distinction between an operator and an impersonator of it.

@@

\end{proof}

[Is this the right math?]

Let $c_\ell$ denote a chaperone installed at the site $\ell$.
Let $P_0(c_\ell)$ denote that $c_\ell$ is well-behaved.
Let $R(c_{\ell_0},c_{\ell_1})$ if $c_{\ell_0}$ depends on $c_{\ell_1}$.
Let $\bar{R}$ be the transitive, reflexive closure of $R$.
A chaperone $c_{\ell_0}$ is dischargeable if $\forall c_\ell$ such that $\bar{R}(c_\ell,c_{\ell_0})$, $P_0(c_\ell)$.
A chaperone-installment site with label $\ell$ is dischargeable if all chaperones $c_{\ell_0}$ are dischargeable for $\ell_0=\ell$.
%For a chaperone-installment site with label $\ell$, let $P_0(\ell)$ hold when $\forall c_\ell

>>>>

We are after a similar result for chaperoned procedures which will allow us to identify meaningful chaperoned values---values which influence the behavior of the program via chaperoning.
While chaperoned values do so primarily by raising an error, chaperoned procedures can also affect program behavior by diverging.
While we expect that programs will seldom result in divergence, and that specifically at the hand of a chaperoned value, specifications expressed in the source language are susceptible to the same programmer errors as the program proper, which includes unintended partiality.

@@

If a specific chaperoned value is not meaningful in a particular execution, it can be soundly \emph{discharged}, meaning that it is safe to replace its particular installation with identity if it is possible to do so.
(The soundness claim does not say it is necessarily possible---only safe---to do so.)
The conditions for a chaperone to be meaningful are
\begin{enumerate}
\item the negative guard causes an error to be raised,
\item \label{con:neg-divergence} the negative guard diverges,
\item the negative guard does not return the same number of results as arguments given (or one extra),
\item the extra result, if provided, is not a procedure,
\item the results of the negative guard are not, at most, chaperones of the arguments,
\item the positive guard does not accept the number of results produced by the chaperoned procedure,
\item the positive guard causes an error to be raised,
\item the positive guard diverges,
\item the positive guard does not return the same number of results as arguments given,
\item the results of the positive guard are not, at most, chaperones of the arguments, or
\item its application effects the installation of a meaningful chaperone.
\end{enumerate}

@@

This strategy immediately shifts our focus from dynamic to static points of installation in the following way:
First, chaperone installation may not be the only operation performed at a call site.
Thus, we must restrict ourselves to call sites dedicated to that purpose.
Second, even once we have restricted ourselves to only chaperone-installation sites, there may be more than one chaperone installed at each site.
Thus, we must restrict ourselves to sites at which no meaningful chaperone is installed.
Finally, we must extend these conditions from a particular execution to all possible executions.

% multiple operations performed at a particular site
% multiple chaperones installed at a particular site (during one particular execution)
% the chaper
%First, multiple chaperone installations may occur at a given application site during program execution, so we cannot, in all circumstances, selectively prevent the creation of a single chaperone.

%to points of installation

%This strategy immediately raises three issues.
%Such a strategy affects all executions

%Using this approach to discharge, it is clear that chaperone installment at a particular site can be discharged only if it meets these conditions for every possible execution.
These conditions seem onerous.
Indeed, in Racket, users typically access chaperone installation through a library wrapper so all chaperones in the entire program share the same installation site.
This, in turn, means that a given chaperone can be safely discharged only if \emph{all} chaperones in all executions of the program can be safely discharged.
%In other words, the program must be correct up to the encoded specification.
Fortunately, with access to the whole program, we may inline definitions which allows us to multiply the number of installment sites in the program.

%This immediately raises two issues.

%Second, chaperone installation may not be the only operation performed by an application site.
%We address both by analyzing application sites and then only those in which \scheme{chap-op} is the only operator, which we refer to as \emph{chaperone-installation sites}.
%The conditions to discharge a given site are stronger than that of a chaperone installed at that site: instead of demonstrating that a single chaperone is not meaningful, we must demonstrate that every chaperone created at that site is not meaningful \emph{and} that the chaperone installations themselves are not meaningful.

@@

%The conditions necessary for the sound discharge of a chaperone are more easily constructed by considering the conditions sufficient for a chaperone to be meaningful.
%If we can enumerate a complete set of these conditions in the sense that any meaningful chaperone must meet at least one of the conditions, we obtain the soundness conditions by complement.

%For a given program execution, if neither a chaperone nor its installation raises an error or diverges, it is a candidate for \emph{discharge}, or the targeted prevention of its installation.
%It is only a candidate because its application may cause a meaningful chaperone to be installed, i.e., a chaperone that is not so well-behaved.
%In this case, discharging the original chaperone would prevent a meaningful chaperone from being created which would make discharge unsound.
%Then, intuitively, a chaperone can be discharged if it will not raise an error or diverge  \emph{and} any chaperones it installs can be discharged.

@@

It may be that a chaperone is not entirely well-behaved but not in any way unique to chaperones.
For instance, consider a program with a single chaperone which provokes an error unspecific to chaperones, such as returning too many results from the positive guard.
Our meaningfulness conditions disqualify this chaperone from discharge, and for good reason---it influences the behavior of the program.
But, if no other conditions object, we can safely \emph{demote} it to a mere impersonator by replacing \scheme{chap-op} with \scheme{imp-op} at its installation site.

In a program with a plurality of related chaperones, we must be careful, however.
%Much of the run-time overhead of chaperones is shared by impersonators.
%Nevertheless, the enforcement of chaperone-specific propertiesinvariants specific to chaperones is somewhat costly---specifically, the enforcement that chaperone guards at most install chaperones on their arguments.
%The conditions to \emph{demote} the set of chaperones installed at a particular site to impersonators (by replacing the \scheme{chap-op} operator with \scheme{imp-op}) are immediately weaker than to discharge the site entirely: we need to demonstrate only that the guards of chaperones installed there do not arbitrarily modify their arguments (conditions 5 and 10 above) for it to be a candidate for demotion.
If one of a site's chaperones is installed on the argument of a chaperone guard, demoting it to install merely an impersonator will make an otherwise well-behaved chaperone not so.
For instance, the program
\begin{schemedisplay}
(let ([sabotage (chap-op (LAMBDA (f) f) (LAMBDA (f) (chap-op f raise)))])
  (let ([g (sabotage (LAMBDA (x) x))])
    (g 1)))
\end{schemedisplay}
defines a chaperoned procedure \scheme{sabotage} which wraps higher-order arguments to fail on application, defines \scheme{g} by applying \scheme{sabotage} to the identity function, and applies \scheme{g} to \scheme{1}, which raises an error.
At no point does \scheme{g} violate chaperone-specific restrictions, so it is tempting to alter its installation site (within the guard of \scheme{sabotage}) to merely impersonate the first argument.
If done, however, the program would fail at the earlier point when \scheme{sabotage} is called which certainly differs in meaning.

This observation provides us with a sufficient condition for sound chaperone demotion: a chaperone may be demoted if it doesn't provoke a chaperone-specific error and if the chaperone which caused its installation, if any, is soundly demoted.
We can lift this condition to chaperone-installation sites in a way similar to the discharge condition.
% neither of its guards do more than install chaperones on their arguments for all program executions and the chaperone which guard installed it, if any, can be and is demoted. 
%We once again lift the conditions to chaperone-installment sites so that a chaperone-installment site can be demoted if all chaperones installed there in any program execution can be demoted and the installment sites of any installing chaperones can be and a re demoted.

@@

We reverse the notion of dependence to say that certain control points depend on the active procedures on the stack.
With our ultimate goal being to prevent chaperone installation, we are only concerned with the chaperone guard procedures depended upon at each point of interest.
Even in the presence of tail-call optimization, as in \chapcalc, the presence of active chaperone guard procedures is always recorded on the stack, and we can apply Harrison's dependence reasoning fairly directly.

@@

%Let $H_c(c_{\ell_1},c_{\ell_0})$ if $c_{\ell_1}$ depends on $c_{\ell_0}$.

@@

%Let $H_s(s_{\ell_1},c_{\ell_0})$ if $s_{\ell_1}$ depends on $c_{\ell_0}$.

@@

%Let $H_e(e,c_\ell)$ if $e$ depends on $c_\ell$.

@@

%Let $H_e(e,s_\ell)$ if $e$ depends on $s_\ell$.

@@

%Let $G(c_\ell)$ hold if $c_\ell$ is well-behaved.

@@

%Transitive, reflexive closure of $H_c$ is $\bar{H_c}$,
%then $c_{\ell_0}$ is dischargeable if for all $c_{\ell_1}$ such that $\bar{H_c}(c_{\ell_1},c_{\ell_0})$, $G(c_{\ell_1})$.

@@

%\begin{proof}
%By induction on the reduction relation.
%Suppose $p\rrs v$.
%Then none of the chaperone guards (or any guard, for that matter) must raise an error.
%Then there is nothing to worry about.
%That is, $D_\ell(p)\rrs v$, where we make no distinction between an operator and a chaperone of it.

%Suppose $p\Uparrow b$. 
%Then an error must be raised while a chaperone (with $b$ derived from its label) is active.
%Then that chaperone depends on the error.
%Then that chaperone-installment site will not be a candidate for deletion.
%It is not dischargeable.
%Then it's not labelled $\ell$.
%Since this is true for every dependent non-dischargeable site (that it's not labelled $\ell$), then $D_\ell(p)\Uparrow b$.

%Suppose $p\Uparrow$.
%If the installation or application of a chaperone could cause divergence, it won't be dischargeable.
%Then the chaperone site labelled $\ell$ must not cause divergence.
%Then $D_\ell(p)\Uparrow$.
%\end{proof}

@@

%\begin{proof}
%Suppose $p\rrs v$.
%Then no error is thrown in regular execution.
%Because we will not demote a site unless there is no possibility that its result is used %by a chaperone, demoting the site will not introduce an error.
%Therefore, $E_\ell(p)\rrs v$ where we make no distinction between an operator and an impersonator of it.

%\end{proof}

@@

%\begin{proof}
%Suppose $p\rrs v$.
%Then no error is thrown in regular execution.
%Because we will not demote a site unless there is no possibility that its result is used by a chaperone, demoting the site will not introduce an error.
%Therefore, $E_\ell(p)\rrs v$ where we make no distinction between an operator and an impersonator of it.

@@

%\end{proof}

%[Is this the right math?]

%Let $c_\ell$ denote a chaperone installed at the site $\ell$.
%Let $P_0(c_\ell)$ denote that $c_\ell$ is well-behaved.
%Let $R(c_{\ell_0},c_{\ell_1})$ if $c_{\ell_0}$ depends on $c_{\ell_1}$.
%Let $\bar{R}$ be the transitive, reflexive closure of $R$.
%A chaperone $c_{\ell_0}$ is dischargeable if $\forall c_\ell$ such that $\bar{R}(c_\ell,c_{\ell_0})$, $P_0(c_\ell)$.
%A chaperone-installment site with label $\ell$ is dischargeable if all chaperones $c_{\ell_0}$ are dischargeable for $\ell_0=\ell$.
%For a chaperone-installment site with label $\ell$, let $P_0(\ell)$ hold when $\forall c_\ell

### 2013-11-16 pldi2014/content/introduction.tex
Everything but discharge section.

The moderation of these costs has been studied extensively.
Herman et al.~\cite{herman2010space} observed the potential accumulation of casts in the blame calculus---similar in principle to the accumulation of guards by our contracted \scheme{fact} procedure---and applied the coercion calculus of Henglein~\cite{henglein1994dynamic} to compact them.
Siek and Wadler~\cite{siek2010threesomes} consider the same problem and apply their own solution of \emph{threesomes}.
Siek et al.~\cite{siek2006gradual} explicate type systems which allow \emph{gradual typing} in which they provide a low-cost path from dynamically- to statically-checked invariants.
Flanagan and Knowles~\cite{flanagan2006hybrid} introduce \emph{hybrid type checking} in which static analysis is used to establish contract compliance statically where possible and insert dynamic checks where not.
%Using this final approach, a disciplined programmer enjoys the early safety guarantees and performance characteristics of statically-verified code or suffers the cost of belated error detection of dynamically-checked specifications.

Unfortunately, modular development can impede these analyses because the greatest amount of knowledge a module can have about its imports \emph{is} their behavioral contract.\footnote{Indeed, it can be less than this, as when the behavioral specification is not codified entirely within the program.}
For example, recognizing that contracts constitute total formal knowledge, Tobin-Hochstadt and Van Horn \cite{tobin2012higher} conduct an intra-module analysis which identifies extra-module references with opaque values that satisfy their contract.
% also, contracts are attached to values that flow across module boundaries.
Nevertheless, static analyses typically perform best when given whole programs.
For instance, the whole-program ML compiler MLton \cite{weeks2006whole} is able to perform thorough inter-procedural and inter-modular analysis and effect significant program optimizations.
We similarly expect the establishment of certain properties to become more accessible as code knowledge increases.
%We expect the emergence of the possibility of more properties being proved about the code as more knowledge about it becomes available.

@@

Recent advances in higher-order program analysis \cite{earl2012introspective} provide perfectly precise return flow and allow the form of the stack to be approximated at each control point in the program.
The original purpose of this approximation was to provide an abstract garbage collector a conservative set of live frames.
We repurpose this capability to analyze a notion of dependence similar to that of Harrison~\cite{harrison1989interprocedural} allowing us to judge the safe prevention of dynamic contract installation---that is, contracts \emph{installed} at run time---\emph{even when the installation of other contracts depends on them}.
The expressiveness of Racket's contract language admits dynamic checks which diverge, so we must integrate a termination analysis based on Sereni's~\cite{sereni2007termination} that informs the discharge of dependently-installed contracts.

We proceed as follows:
We discuss interposition and its relationship to behavioral contracts in more depth in Section 2.
We introduce the syntax and semantics of \chapcalc, a core calculus with facilities for interposing higher-order values, in Section 3.
We present two theorems which establish the conditions for the safe removal or weakening of interposition in Section 4.
We present a static analysis which can judge when these conditions hold in Section 5.
We discuss the influence of parameters of the analysis on precision and performance in Section 6.
%We discuss some Racket language features which require more care to integrate with the present analysis in Section 7.
We discuss related work in Section 7.
We conclude by discussing the integration of some Racket language features with this analysis in Section 8.
%future application and direction of this work in Section 8.

>>>>

The moderation of these costs has been studied extensively.
Herman et al.~\cite{herman2010space} observed the potential accumulation of casts in the blame calculus---similar in principle to the accumulation of guards by our contracted \scheme{fact} procedure---and applied the coercion calculus of Henglein~\cite{henglein1994dynamic} to compact them.
Siek and Wadler~\cite{siek2010threesomes} consider the same problem and apply their own solution of \emph{threesomes}.
Siek et al.~\cite{siek2006gradual} explicate type systems which allow \emph{gradual typing} in which they provide a low-cost path from dynamically- to statically-checked invariants.

The approach we adopt is most similar to Flanagan and Knowles'~\cite{flanagan2006hybrid} \emph{hybrid type checking} in which static analysis is used to establish contract compliance statically where possible and insert dynamic checks where not.
%Using this final approach, a disciplined programmer enjoys the early safety guarantees and performance characteristics of statically-verified code or suffers the cost of belated error detection of dynamically-checked specifications.
Unfortunately, modular development can impede these analyses because the greatest amount of knowledge a module can have about its imports \emph{is} their behavioral contract.\footnote{Indeed, it can be less than this, as when the behavioral specification is not codified entirely within the program.}
For example, recognizing that contracts constitute total formal knowledge, Tobin-Hochstadt and Van Horn \cite{tobin2012higher} conduct an intra-module analysis which identifies extra-module references with opaque values that satisfy their contract.
% also, contracts are attached to values that flow across module boundaries.
Nevertheless, static analyses typically perform best when given whole programs.
For instance, the whole-program ML compiler MLton \cite{weeks2006whole} is able to perform thorough inter-procedural and inter-modular analysis and effect significant program optimizations.
We similarly expect the establishment of certain properties to become more accessible as code knowledge increases.
%We expect the emergence of the possibility of more properties being proved about the code as more knowledge about it becomes available.

@@

Recent advances in higher-order program analysis \cite{earl2012introspective} provide perfectly precise return flow and allow the form of the stack to be approximated at each control point in the program.
The original purpose of this approximation was to provide an abstract garbage collector a conservative set of live frames.
We repurpose this capability to analyze a notion of dependence similar to that of Harrison~\cite{harrison1989interprocedural} allowing us to judge the safe prevention of dynamic contract installation---that is, contracts \emph{installed} at run time---\emph{even when the installation of other contracts depends on them}.
The expressiveness of Racket's contract language admits partial dynamic checks, so we must integrate a termination analysis based on Sereni's~\cite{sereni2007termination} that informs the discharge of dependently-installed contracts.

We proceed as follows:
We discuss interposition and its relationship to behavioral contracts in more depth in Section 2.
We introduce the syntax and semantics of \chapcalc, a core calculus with facilities for interposing higher-order values, in Section 3.
We present two theorems which establish the conditions for the safe removal or weakening of interposition in Section 4.
We present a static analysis which can judge when these conditions hold in Section 5.
We discuss the influence of parameters of the analysis on precision and performance in Section 6.
%We discuss some Racket language features which require more care to integrate with the present analysis in Section 7.
%We discuss related work in Section 7.
We discuss the integration of other Racket language features with this analysis in Section~\ref{sec:limitations}.
We conclude with future application and direction of this work in Section~\ref{sec:conclusion}.

### 2013-11-16 pldi2014/content/limitations.tex
Everything but discharge section.

Our calculus has an impoverished value domain, so our termination judgments of guards are nearly complete in practice as a result of only a simple flow analysis.
%Because, in practice, guards are not recursive themselves or any procedures in the call graph. Essentially, call graphs for guard procedures are trees.)
Contracts on compound data complicate matters because the data may be defined inductively which makes recursive checking natural.
We can accomodate this in particular by integrating size-change analysis~\cite{lee2001size} into our abstract interpretation.
This would be novel as higher-order size-change analysis~\cite{sereni2007termination} has heretofore been applied only to variants of $k$-CFA~\cite{shivers1991control}.

@@

@@

Other features inherited from Scheme are mutation, variable-arity procedures.

On its face, the integration of an error-handling construct is tractable, at least with the current, limited calculus.
We simply need to keep in mind that errors influence the behavior of the program whether they are redirected or not.
Essentially, we would treat exception handlers for error-dependence analysis as if they didn't exist.
%mutation
%different types of impersonators
%impersonator properties

@@

In Racket, the \scheme{chaperone-of?} relation is accessible through a predicate procedure.
However, its presence in Racket has different implications than its introduction to \chapcalc, which can use it to determine whether two procedures are intensionally equivalent.
(In Racket, its use in that fashion is restricted to determining whether two references indicate the same procedure.)

Chaperones are intended to be treated like the original value.
The \scheme{chaperone?} predicate is designated as the single way to detect chaperones at runtime, though its use is discouraged.

>>>>

\subsection{Rest parameters}

As it stands, our analysis depends on procedure arguments being represented statically.
Typical Racket programs frequently use \emph{rest} parameters to dynamically gather provided arguments and \scheme{apply} to dynamically provide arguments to a procedure.

\subsection{Inductively-defined values}

Our calculus has an impoverished value domain, so our termination judgments of guards are nearly complete in practice as a result of only a simple flow analysis.
%Because, in practice, guards are not recursive themselves or any procedures in the call graph. Essentially, call graphs for guard procedures are trees.)
Contracts on compound data complicate matters because the data may be defined inductively which makes recursive checking natural.
We can accomodate this in particular by integrating size-change analysis~\cite{lee2001size} into our abstract interpretation.
This would be novel as higher-order size-change analysis~\cite{sereni2007termination} has heretofore been applied only to variants of $k$-CFA~\cite{shivers1991control}.
As seen with chaperones, these types of values can destroy precision when determining whether a chaperone guard is well-behaved.

\subsection{Mutation}

Mutation compounds the problem of static determination of the $\mathrm{chaperone\mhyphen of?}$ relation.

\subsection{Impersonator properties}

@@

\subsection{Other impersonator types}

@@

%Other features inherited from Scheme are mutation, variable-arity procedures.

\subsection{Exception handlers}

On its face, the integration of exception handlers is tractable, at least in the current, limited calculus.
We simply need to keep in mind that errors influence the behavior of the program whether they are redirected or not.
Essentially, we would treat exception handlers for error-dependence analysis as if they didn't exist.
%mutation
%different types of impersonators
%impersonator properties

@@

\subsection{Chaperone-identifying predicates}

In Racket, the \scheme{chaperone-of?} relation is accessible through a predicate procedure.
We omitted it from \chapcalc\ for simplicity, but programmer access does not seem to complicate our analysis.
%However, its presence in Racket has different implications than its introduction to \chapcalc, which can use it to determine whether two procedures are intensionally equivalent.
%(In Racket, its use in that fashion is restricted to determining whether two references indicate the same procedure.)
Racket also provides a \scheme{chaperone?} predicate which discriminates chaperoned values from all others.
We might be able to handle this predicate by tainting the abstract chaperone value to which it is applied.
In practice, this probably would not disqualify any otherwise-qualified candidates for discharge since chaperones are intended to be treated transparently and use of \scheme{chaperone?} is discouraged.
